*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-04 Code/ManuscriptPURFits.Rmd
Added letters to PUR figures

```{r Pur plot strain faceted in columns}

@@

        legend.spacing.x = unit(0.75, 'cm')) +

@@

      )

@@

# This labels each pane with a letter but turns of the dacet labels
# PhotoperiodPurFitsPlots <- tag_facet(PhotoperiodPurFitsPlots, open = "", close = "", tag_pool = LETTERS, size = 8)

Would like to add Compact Letter Displays (cld) but do not know how. Some examples:
https://stackoverflow.com/questions/70854552/any-other-options-besides-the-traditional-cld-bar-graph/70863531#70863531
https://stackoverflow.com/questions/72756490/loop-tukey-post-hoc-letters-extraction/73274274#73274274

@@

ggsave(file = file.path(PlotsOut, paste("PhotoperiodPurFitsPlots",".tiff",sep = "")), plot = PhotoperiodPurFitsPlots, device = NULL, scale = 1, height=30, width= 44, units = c("cm"),dpi = 300, limitsize = TRUE)

@@

```{r Pur plot strain faceted in columns}

@@

AllFitsCombinedWL |>
  geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 2, alpha = 0.7) +
    geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 1.0, alpha = 0.5) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1.0, alpha = 0.3, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "MED4")) + 
   geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_450) %>%  filter(O2_uM == 2.5, Strain == "MED4",  WL == 450)) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_660) %>%  filter(Strain == "MED4", WL == 660))+
    geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_WW) %>%  filter(Strain == "MED4", WL == "WW")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1.0, alpha = 0.3, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "SS120")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6,  data = . %>% unnest(Predict_450) %>%  filter(Strain == "SS120", O2_uM != 25, WL == 450)) +
 geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6,  data = . %>% unnest(Predict_WW) %>%  filter(Strain == "SS120", O2_uM != 250, WL == "WW")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1.0, alpha = 0.3, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "MIT9313")) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_660) %>%  filter(Strain == "MIT9313", O2_uM != 250, WL == 660))+
  geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6,  data = . %>% unnest(Predict_WW) %>%  filter(Strain == "MIT9313", O2_uM == 2.5, WL == "WW")) +
 labs(title = "Anova fit comparisons of WL to pooled", y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "PUR (µmol photons" ~m^-2 ~d^-1~")") +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", 660, 450)) +
  scale_shape_manual(values = PhotoperiodShapes)  +
 guides(linetype = guide_legend(nrow = 1),
        fill = "none",
        color = guide_legend(nrow = 1)) +
 guides(shape = guide_legend(nrow = 1)) +
        legend.key.size = unit(0.1, 'cm'), #change legend key size
        legend.key.height = unit(0.1, 'cm'), #change legend key height
        legend.key.width = unit(0.1, 'cm'), #change legend key width
        legend.title =  element_text(size = 25),  #change legend title font size
        legend.text = element_text(size = 25),   #change legend text font size
        legend.position = "top") +
  labs( shape = "Photoperiod (h)",
        color = "Growth Waveband (nm)",
       linetype = "Photoperiod lines"
      )

@@

      )

>>>>

```{r Pur plot strain faceted in columns, fig.height=10, fig.width=10}

@@

        legend.spacing.x = unit(0.75, 'cm'),
        tagger.panel.tag.background = element_blank()) +

@@

      ) +
  tagger::tag_facets(tag_levels  = c("A", "I"), tag_suffix = "")

@@

# This labels each pane with a letter but turns of the dacet labels
# PhotoperiodPurFitsPlots <- tag_facet(PhotoperiodPurFitsPlots, open = "", close = "", tag_pool = LETTERS, size = 8)
# 
# PhotoperiodPurFitsPlots
```

Would like to add Compact Letter Displays (cld). Some examples:
https://stackoverflow.com/questions/70854552/any-other-options-besides-the-traditional-cld-bar-graph/70863531#70863531
https://stackoverflow.com/questions/72756490/loop-tukey-post-hoc-letters-extraction/73274274#73274274

@@

# ggsave(file = file.path(PlotsOut, paste("PhotoperiodPurFitsPlots",".tiff",sep = "")), plot = PhotoperiodPurFitsPlots, device = NULL, scale = 1, height=30, width= 44, units = c("cm"),dpi = 300, limitsize = TRUE)

@@

```{r Pur plot strain faceted in columns , fig.height=10, fig.width=10}

@@

# AllFitsCombinedWL |>
#          WL = factor(Data_all_WL, levels = c("WW", 660, 450)),
#          Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313"))) %>%
#   geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 2, alpha = 0.7) +
#     geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 1.0, alpha = 0.5) +
#   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1.0, alpha = 0.3, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "MED4")) + 
#    geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_450) %>%  filter(O2_uM == 2.5, Strain == "MED4",  WL == 450)) +
#   geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_660) %>%  filter(Strain == "MED4", WL == 660))+
#     geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_WW) %>%  filter(Strain == "MED4", WL == "WW")) +
#    geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1.0, alpha = 0.3, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "SS120")) +
#    geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6,  data = . %>% unnest(Predict_450) %>%  filter(Strain == "SS120", O2_uM != 25, WL == 450)) +
#  geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6,  data = . %>% unnest(Predict_WW) %>%  filter(Strain == "SS120", O2_uM != 250, WL == "WW")) +
#    geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1.0, alpha = 0.3, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "MIT9313")) +
#   geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6, data = . %>% unnest(Predict_660) %>%  filter(Strain == "MIT9313", O2_uM != 250, WL == 660))+
#   geom_line(aes(x = PURPhotonDose_day, y = .fitted, color = as.factor(WL)), size = 0.6,  data = . %>% unnest(Predict_WW) %>%  filter(Strain == "MIT9313", O2_uM == 2.5, WL == "WW")) +
#  labs(title = "Anova fit comparisons of WL to pooled", y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "PUR (µmol photons" ~m^-2 ~d^-1~")") +
#   # scale_linetype_manual(values = MyLineTypes, labels = c("16 hr", "12 hr", "8 hr", "4 hr")) +
#   ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), cols = vars(Strain), 
#                     labeller = labeller(Strain = label_value, O2_uM = label_value, OxygenLabel = label_value)) +
#    theme( strip.text = element_text(size = 25),
#         axis.title.y = element_text(color = "black", size = 30,  hjust = 0.40),
#         axis.title.x = element_text(color = "black", size = 30),
#         axis.text.x = element_text(size = 20, color = "black", angle = 45, hjust = 1.0),
#         axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
#         # strip.background=element_rect(color="grey87", fill = "grey82"),
#         legend.key.size = unit(0.1, 'cm'), #change legend key size
#         legend.key.height = unit(0.1, 'cm'), #change legend key height
#         legend.key.width = unit(0.1, 'cm'), #change legend key width
#         legend.title =  element_text(size = 25),  #change legend title font size
#         legend.text = element_text(size = 25),   #change legend text font size

@@

      ) +
  tagger::tag_facets(tag_levels  = c("A", "I"), tag_suffix = "")

### 2024-03-04 Code/Manuscript_O2.Rmd
Added letters to PUR figures

knitr::write_bib(c("knitr", "tidyverse", "lubridate", "ggplot2", "broom", "grid", "bookdown", "glue", "cowplot", "ggpubr", "ggh4x", "png", "Studio", "zoo", "photobiology", "photobiologyWavebands", "data.table", "googlesheets4",  "mgcv", "minpack.lm",  "egg", 'RPackages.bib'))

@@

We analyzed  growth rates to determine the viability of *P. marinus* MED4, a clade HLI ecotype found near the ocean surface,
*P. marinus* SS120, a clade LLII/III ecotype found deep in the water column, and *P. marinus* MIT9313, a clade LLIV ecotype also found in deep oceans including OMZ, under a matrix of spectral, light level, photoperiod and oxygen conditions approxmating present day and hypothetical future niches. Low light levels in the blue spectral waveband prevail in deep ocean waters [@morelAvailableUsableStored1978; @holtropVibrationalModesWater2021], while solar incidence angle and light attenuation also progressively attenuate the photoperiod with depth beneath the ocean surface.

*Prochlorococcus marinus* MED4 has a physiological requirement for more than 4 h of light per day; thus this strain will not exploit habitats with short photoperiods, typical of temperate winter or light attenuated depths, even if water temperature warms into the clade LLII/III tolerance range. MED4 is also excluded from the lowest oxygen habitats, represented by our 2.5 µM O~2~ experimental conditions. MED4 can, however, grow under OMZ regions with slightly higher [O~2~], as demonstrated by the growth under our 25 µM O~2~ experiments. Genomic analyses (Figure \@ref(fig:ProchlorococcusEnzymeKms))  and previous transcriptional analyses [@bergResponsesPsbAHli2011]  suggest MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2~] environments. Photoinactivation, does, however, impose an increased cost of growth upon MED4, since growth under red light, to lower the rate of photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under blue light. TARA Oceans Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. Delmont and Eren [@delmontLinkingPangenomesMetagenomes2018] did not analyze data from depths beyond the subsurface chlorophyll maximum layer, the layer in the water column where the chlorophyll *a* concentration peaks, nor did they report the [O~2~] at those depths. Our growth findings are also consistent with Figure\@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, clade HLI, at depths up to 200 meters with O~2~ of 15 µM.

@@

XXX PLOS requires grant numbers XXX
# Funding  {.unnumbered}
To Be Entered through PLoSONe system and deleted here
Czech Academy of Science visiting fellowship supporting DAC work at AlgaTech (OP)
Canada Research Chair in Phytoplankton Ecophysiology (DAC)
Natural Sciences and Engineering Research Council of Canada, 'Latitude and Light' (DAC)
Canada Foundation for Innovation (DAC)
New Brunswick Foundation for Innovation (DAC)
Rice Graduate Fellowship (MS)

>>>>

knitr::write_bib(c("knitr", "tidyverse", "lubridate", "ggplot2", "broom", "grid", "bookdown", "glue", "cowplot", "ggpubr", "ggh4x", "png", "Studio", "zoo", "photobiology", "photobiologyWavebands", "data.table", "googlesheets4",  "mgcv", "minpack.lm",  "tagger", 'RPackages.bib'))

@@

We analyzed  growth rates to determine the viability of *P. marinus* MED4, a clade HLI ecotype found near the ocean surface,
*P. marinus* SS120, a clade LLII/III ecotype found deep in the water column, and *P. marinus* MIT9313, a clade LLIV ecotype also found in deep oceans including OMZ, under a matrix of spectral, light level, photoperiod and oxygen conditions approximating present day and hypothetical future niches. Low light levels in the blue spectral waveband prevail in deep ocean waters [@morelAvailableUsableStored1978; @holtropVibrationalModesWater2021], while solar incidence angle and light attenuation also progressively attenuate the photoperiod with depth beneath the ocean surface.

*Prochlorococcus marinus* MED4 has a physiological requirement for more than 4 h of light per day; thus this strain will not exploit habitats with short photoperiods, typical of temperate winter or light attenuated depths, even if water temperature warms into the clade HLI tolerance range. MED4 is also excluded from the lowest oxygen habitats, represented by our 2.5 µM O~2~ experimental conditions. MED4 can, however, grow under OMZ regions with slightly higher [O~2~], as demonstrated by the growth under our 25 µM O~2~ experiments. Genomic analyses (Figure \@ref(fig:ProchlorococcusEnzymeKms))  and previous transcriptional analyses [@bergResponsesPsbAHli2011]  suggest MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2~] environments. Photoinactivation, does, however, impose an increased cost of growth upon MED4, since growth under red light, to lower the rate of photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under blue light. TARA Oceans Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. Delmont and Eren [@delmontLinkingPangenomesMetagenomes2018] did not analyze data from depths beyond the subsurface chlorophyll maximum layer, the layer in the water column where the chlorophyll *a* concentration peaks, nor did they report the [O~2~] at those depths. Our growth findings are also consistent with Figure\@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, clade HLI, at depths up to 200 meters with O~2~ of 15 µM.

@@

XXX PLOS requires grant numbers XXX

# Funding  {.unnumbered}

To Be Entered through PLoSONe system and deleted here
Czech Academy of Science visiting fellowship supporting DAC work at AlgaTech (OP)
Canada Research Chair in Phytoplankton Ecophysiology (DAC)
Natural Sciences and Engineering Research Council of Canada, 'Latitude and Light' (DAC)
Canada Foundation for Innovation (DAC)
New Brunswick Foundation for Innovation (DAC)
Rice Graduate Fellowship (MS)
